---
title: "Generate CartoDB Data"
output: html_document
---

```{r loadData}
source("./dataCleaning.R")
source("./dataAnalysis.R")
set.seed(58978160)

dataDirectory <- "./Data"
analyticDataPath <- "./AnalyticData"
maximumPercentMissing <- 10.0

if (is.na(file.info(dataDirectory)[1,"isdir"])) {
    downloadData(dataDirectory)
}

if (is.na(file.info(file.path(analyticDataPath,
                              "/trainingData.csv"))[1,"size"])) {
    createAnalyticDataset(dataDirectory,
                          maximumPercentMissing,
                          analyticDataPath)
}
```

```{r createCartoDBData, echo=FALSE}
zipCodeStatisticsFile <- file.path(analyticDataPath,
                                   "zipCodeLevelStatistics.RData")

if (is.na(file.info(zipCodeStatisticsFile)[1,"size"])) {
    trainingData <- loadAnalyticData(analyticDataPath,
                                     "trainingData.csv")

    summaryIssueCategoryPercentage <-
        initializeSummaryIssueCategoryPercentage(trainingData)

    issueCategories <- summaryIssueCategoryPercentage[1:5,
                                                      c("issuecategory")]

    rowsToSelect <- vector()
    for (curIssueCategory in issueCategories) {
        rowsToSelect <- append(rowsToSelect,
                               which(trainingData$issuecategory ==
                                     curIssueCategory))
    }
    rowsToSelect <- sort(rowsToSelect)

    segmentedData <- trainingData[rowsToSelect,]

    segmentedData$issuecategory <-
        as.factor(as.character(segmentedData$issuecategory))

    stateAbbreviation <- unique(segmentedData$state)

    zipCodeIssueCategoryPercentage <- 
        initializeZipCodeIssueCategoryPercentage(segmentedData,
                                                 stateAbbreviation)

    save(list=c('zipCodeIssueCategoryPercentage',
                'issueCategories'), file=zipCodeStatisticsFile)
#-------------------------------------------------------------
} else {
    load(file=zipCodeStatisticsFile)
}

for (categoryIdx in seq_len(length(issueCategories))) {
    curIssueCategoryRows <-
        which(zipCodeIssueCategoryPercentage$issuecateogry ==
              issueCategories[categoryIdx])

    curPercentageData <- 
        zipCodeIssueCategoryPercentage[curIssueCategoryRows,]

    variableNames <- colnames(curPercentageData)
    
    variableNames <- gsub("percentage",
                          paste0(issueCategories[categoryIdx],
                                 "percentage"),
                          variableNames)
    
    colnames(curPercentageData) <- variableNames

    curPercentageData <-
        curPercentageData[,variableNames[!grepl("^issuecateogry$",
                                                variableNames)]]

    # http://stackoverflow.com/questions/2232699/
    #   how-to-do-a-data-table-merge-operation
    if (categoryIdx == 1) {
        cartoDBData <- curPercentageData
    #--------------------------------------------------
    } else {
        cartoDBData <- merge(cartoDBData,
                             curPercentageData,
                             by="zipcode",
                             all.x=TRUE,
                             all.y=TRUE)

        variables <- colnames(cartoDBData)
        yVariableRegex <- "^[a-z]+\\.y$"
        xMissingRows <- which(is.na(cartoDBData$city.x))

        if (length(xMissingRows) > 0) {
            variablesToRead <- variables[grepl(yVariableRegex,variables)]
            variablesToWrite <- gsub("\\.y","\\.x",variablesToRead)

            for (idx in seq_len(length(variablesToRead))) {
                cartoDBData[xMissingRows, variablesToWrite[idx]] <-
                    cartoDBData[xMissingRows, variablesToRead[idx]]
            }
        }
        
        cartoDBData <- cartoDBData[,variables[!grepl(yVariableRegex,
                                                    variables)]]

        variables <- colnames(cartoDBData)
        colnames(cartoDBData) <- sub("^([a-z]+)\\.x","\\1",variables)
    }
}

variables <- colnames(cartoDBData)
for (percentageVariable in variables[grepl("^[a-z]+percentage$",
                                           variables)]) {
    naRows <- which(is.na(cartoDBData[,percentageVariable]))
    cartoDBData[naRows,percentageVariable] <- 0.0
}

write.csv(cartoDBData,
          file.path(analyticDataPath,"percentageData.csv"),
          row.names=FALSE)
